### 2012-03-27 doc/theses/mikula/combination_with_existing_methods_of_inference.tex
Master thesis: In progress.
Master thesis: Corrections.

The combination with such methods of inference is straightforward. The rules of the grammar describe a XML structure using the most general aspect of elements and their subelements. Since the statements inferred by our method do not involve the XML structure by defining a subelement structure of elements, there are no conflicts between the grammar rules and our statements that need to be resolved.

@@

In all tree cases, we want to determine elements from the grammar (or their attributes), targeted by the respective paths ($p$ from the first form and $C$ from the second and third form). This is done in two steps. The first one is normalization of a particular PathType and the second one is selection of the targeted elements.

@@

It iterate through steps of a path, maintaining a so-called context set, which is a set of elements to evaluate the current step upon. The evaluation of one step is shown in Algorithm \ref{FUNC_evaluate_step}. For the retaining of readability of the code, we present an evaluation of self or descendant, child, and attribute axes, and nodes specified by name.

@@

At first, the algorithm determine the axis of the step. If it is self or descendant axis, it returns the result of \texttt{evaluateStep\_selfOrDescendant} function. This function returns the given context set extended by all descendants of the elements from the context set in the given grammar.

@@

\section{Saving the Inferred Statements}
In the case of keys, the algorithm iterates through the inferred key statements. A key's context path $C$ is evaluated and the key is assigned to the target elements selected by the context path, one element can be assigned with multiple keys. Additionally, each key is assigned with a list of foreign keys that are referencing it.

In the case of inferred types, the situation is slightly less simple, because there can occur conflicting statements. The examples of such conflicts for path $p$ without predicates are $p_1 \rightarrow date$, $p_2 \rightarrow string$, and, $p_1 \rightarrow byte$, $p_2 \rightarrow int$, where $p_1$, $p_2$ are paths that when stripped of predicates, they equal $p$.

@@

For each PathType $p$ that occurs in the inferred type statements $\mathcal{S}_t$, we find set $T_p$ of all inferred types. $T_p = \{t|(p \rightarrow t) \in \mathcal{S}_t\}$. Then, we create sequence $T_p'$ by ordering the set $T_p$ from the most specific type to the most general one. For example, if $T_p = \{$\texttt{double}, \texttt{byte}, \texttt{int}$\}$, $T_p' = ($\texttt{byte}, \texttt{int}, \texttt{double}$)$.

Since we have the XML data and path $p$ is XQuery path, we can use an XQuery process (a program that evaluates XQuery paths or queries) to select nodes $N$ targeted by $p$. The verification algorithms iterates through $T_p'$ and for each $t' \in T_p'$ it checks if every node in $N$ conforms to $t'$. If so, $t'$ is the inferred type for nodes $N$ (and PathType $p$), else the inferred type is \texttt{string}.

>>>>

The combination with such methods of inference is straightforward. The rules of the grammar describe the XML structure using the most general aspect of elements and their subelements. Since the statements inferred by our method do not involve the XML structure by defining a subelement structure of elements, there are no conflicts between the grammar rules and our statements that need to be resolved.

@@

In all three cases, we want to determine elements from the grammar (or their attributes), targeted by the respective paths ($p$ from the first form and $C$ from the second and third form). This is done in two steps. The first one is normalization of a particular PathType and the second one is selection of the targeted elements.

@@

It iterates through steps of a path, maintaining a so-called \emph{context set}, which is a set of elements to evaluate the current step upon. The evaluation of one step is shown in Algorithm \ref{FUNC_evaluate_step}. For the retaining of readability of the code, we present an evaluation of self or descendant, child, and attribute axes, and nodes specified by name.

@@

At first, the algorithm determines the axis of the step. If it is self or descendant axis, it returns the result of \texttt{evaluateStep\_selfOrDescendant} function. This function returns the given context set extended by all descendants of the elements from the context set in the given grammar.

@@

\section{Saving the Inferred Statements}
In case of keys, the algorithm iterates through the inferred key statements. A key's context path $C$ is evaluated and the key is assigned to the target elements selected by the context path, one element can be assigned with multiple keys. Additionally, each key is assigned with a list of foreign keys that are referencing it.

In case of inferred types, the situation is slightly less simple, because there can occur conflicting statements. The examples of such conflicts for path $p$ without predicates are $p_1 \rightarrow date$, $p_2 \rightarrow string$, and, $p_1 \rightarrow byte$, $p_2 \rightarrow int$, where $p_1$, $p_2$ are paths that when stripped of predicates, they equal $p$.

@@

For each normalized PathType $p$ from the inferred type statements $\mathcal{S}_t$, we find set $T_p$ of all inferred types. $T_p = \{t|(p \rightarrow t) \in \mathcal{S}_t\}$. Then, we create sequence $T_p'$ by ordering the set $T_p$ from the most specific type to the most general one. For example, if $T_p = \{$\texttt{double}, \texttt{byte}, \texttt{int}$\}$, $T_p' = ($\texttt{byte}, \texttt{int}, \texttt{double}$)$.

Since we have the XML data and path $p$ is an XQuery path, we can use an XQuery processor (a program that evaluates XQuery paths or queries) to select nodes $N$ targeted by $p$. The verification algorithm iterates through $T_p'$ and for each $t' \in T_p'$ it checks if every node in $N$ conforms to $t'$. If so, $t'$ is the inferred type for nodes $N$ (and PathType $p$), else the inferred type is \texttt{string}.

### 2012-03-27 doc/theses/mikula/conclusion.tex
Master thesis: Corrections.

\chapter{Conclusion}
The aim of this thesis was to employ XML operations in the XML schema inference. We analysed several existing methods of the XML schema inference and we searched for methods that utilize some XML operations, only to found one, which infers keys from XQuery queries.

Since the notion XML operations is very general and the range of XML technologies is very large, for purpose of this work, we decided to focus on XQuery technology. We made the overview of possible utilization of XQuery queries in the process of XML schema inference.

@@

We also implemented several ideas from the overview to infer XSD built-in types of elements and attributes. And we slightly extended the one existing method dealing with the inference of keys. We experimentally demonstrated that on some input files, results of the extended method are better than results of the original method. However, we did not succeed in a search of an ideal, large enough set of test data. The testing was performed on a small set of input queries and further testing and algorithm tuning is required.

At last, we proposed a simple way how to combine the inferred statements with existing grammar inferring methods of the XML schema inference and we implemented it using jInfer framework. Thus, we created the first complete, ready-to-use, and extensible implementation of the XML xchema inference exploiting XQuery queries besides XML data.

@@

Chapter \ref{CHAPTER_precreation_of_algorithm} discusses some possible future enhancements, as well. A research on possibilities of modifying the grammar rules based on information extracted from XQuery queries may bring interesting results. We analyse the queries statically only, we do not evaluate them. An analysis of queries together with their results can be a topic of some future research.

The utilization of XQuery queries certainly provides a space for incorporating an interaction with user. For example, a user may influence the scoring of inferred keys to get a more precise results.

@@

Besides the mentioned, our opinion is that the most urgent future work is obtaining a large enough test data set, performing proper experiments, and refining the algorithm by modification of its settings (statements inferred from join pattern occurrences, weights, etc.) according to experimental results. This process was slightly suggested in Chapter \ref{CHAPTER_experiments}.

>>>>

\chapter{Conclusion}
The aim of this thesis was to employ XML operations in the XML schema inference. We analyzed several existing methods of the XML schema inference and we searched for methods that utilize some XML operations, only to found one\todo{Jiny vyraz}, which infers keys from XQuery queries.

Since the notion of XML operations is very general and the range of XML technologies is very large, for the purpose of this work, we decided to focus on XQuery technology. We made the overview of possible utilization of XQuery queries in the process of XML schema inference.

@@

We also implemented several ideas from the overview to infer XSD built-in types of elements and attributes. And we extended the one existing method dealing with the inference of keys. We experimentally demonstrated that on some input files, results of the extended method are better than results of the original method. However, we did not succeed in a search of an ideal, large enough set of test data. The testing was performed on a small set of input queries and further testing and algorithm tuning is required.

Finally, we proposed a simple way how to combine the inferred statements with existing grammar inferring methods of the XML schema inference and we implemented it using the jInfer framework. Thus, we created the first complete, ready-to-use, and extensible implementation of the XML xchema inference exploiting XQuery queries besides XML data.

@@

Chapter \ref{CHAPTER_precreation_of_algorithm} discusses some possible future enhancements as well. A research on possibilities of modifying the grammar rules based on information extracted from XQuery queries may bring interesting results. We analyse the queries statically only, we do not evaluate them. An analysis of queries together with their results can be a topic of another future research direction.

The utilization of XQuery queries certainly provides a space for incorporating an interaction with user. For example, a user may influence the scoring of inferred keys to get more precise results.

@@

Besides the mentioned, our opinion is that the most urgent future work is obtaining a large enough test data set, performing proper experiments, and refining the algorithm by modification of its settings (statements inferred from join pattern occurrences, weights, etc.) according to experimental results. This process was suggested in Chapter \ref{CHAPTER_experiments}.

### 2012-03-27 doc/theses/mikula/introduction.tex
Master thesis: Corrections.

With XML used to represent data themselves, it is often needed or convenient to somehow specify a structure of the represented data, their format, internal relations or restrictions, etc. In order to this need, several so-called XML schema languages (XML schemata) have been created. The most used of them are Document Type Definition (DTD) \cite{Bray:08:EML} and XML Schema (or XSD as XML Schema Definition) \cite{Walmsley:04:XSP, Thompson:04:XSP, Malhotra:04:XSP} proposed by the World Wide Web Consorcium (W3C).

Despite an conveyed encouragement to use an XML schema along with an XML data representation, in practise it is done sparsely. Commonly, XML documents are not assigned with their respective XML schema at all or the schema is outdated due to modifications done to a structure of data without updating the schema \cite{Mlynkova:2008:AAX:1494650.1495496}.

>>>>

With XML used to represent data themselves, it is often needed or convenient to somehow specify the structure of the represented data, their format, internal relations or restrictions, etc. In order to this need, several so-called XML schema languages (XML schemata) have been created. The most used of them are Document Type Definition (DTD) \cite{Bray:08:EML} and XML Schema (or XSD as XML Schema Definition) \cite{Walmsley:04:XSP, Thompson:04:XSP, Malhotra:04:XSP} proposed by the World Wide Web Consorcium (W3C).

Despite a conveyed encouragement to use an XML schema along with an XML data representation, in practise it is done sparsely. Commonly, XML documents are not assigned with their respective XML schema at all or the schema is outdated due to modifications done to a structure of data without updating the schema \cite{Mlynkova:2008:AAX:1494650.1495496}.
